---
title: Notes：Converting data between wide and long format
author: Xuejun Liu
date: '2018-02-25'
slug: notes-converting-data-between-wide-and-long-format
categories: ["编程"]
tags: ["R语言","tidyr","reshape2"]
banner: "img/banners/4.jpg"
---

# 1.Problem  

You want to do convert data from a wide format to a long format.

Many functions in R expect data to be in a long format rather than a wide format. Programs like SPSS, however, often use wide-formatted data.



# 2.Solution
There are two sets of methods that are explained below:

* gather() and spread() from the tidyr package. This is a newer interface to the reshape2 package.
* melt() and dcast() from the reshape2 package.

There are a number of other methods which aren’t covered here, since they are not as easy to use:

* The reshape() function, which is confusingly not part of the reshape2 package; it is part of the base install of R.
* stack() and unstack()  

# 3.Sample data

These data frames hold the same data, but in wide and long formats. They will each be converted to the other format below.  
```{r}
olddata_wide <- read.table(header=TRUE, text='
 subject sex control cond1 cond2
       1   M     7.9  12.3  10.7
       2   F     6.3  10.6  11.1
       3   F     9.5  13.1  13.8
       4   M    11.5  13.4  12.9
')
# Make sure the subject column is a factor
olddata_wide$subject <- factor(olddata_wide$subject)
```

```{r}
olddata_long <- read.table(header=TRUE, text='
 subject sex condition measurement
       1   M   control         7.9
       1   M     cond1        12.3
       1   M     cond2        10.7
       2   F   control         6.3
       2   F     cond1        10.6
       2   F     cond2        11.1
       3   F   control         9.5
       3   F     cond1        13.1
       3   F     cond2        13.8
       4   M   control        11.5
       4   M     cond1        13.4
       4   M     cond2        12.9
')
# Make sure the subject column is a factor
olddata_long$subject <- factor(olddata_long$subject)
```

# 4.tidyr

### 4.1 From wide to long
Use ```gather```:
```{r}
olddata_wide
library(tidyr)

# The arguments to gather():
# - data: Data object
# - key: Name of new key column (made from names of data columns)
# - value: Name of new value column
# - ...: Names of source columns that contain values
# - factor_key: Treat the new key column as a factor (instead of character vector)
data_long <- gather(olddata_wide, condition, measurement, control:cond2, factor_key=TRUE)
data_long
```

In this example, the source columns that are gathered are specified with control:cond2. This means to use all the columns, positionally, between control and cond2. Another way of doing it is to name the columns individually, as in:

```{r}
gather(olddata_wide, condition, measurement, control, cond1, cond2)
```

If you need to use gather() programmatically, you may need to use variables containing column names. To do this, you should use the gather_() function instead, which takes strings instead of bare (unquoted) column names.

```{r}
keycol <- "condition"
valuecol <- "measurement"
gathercols <- c("control", "cond1", "cond2")

gather_(olddata_wide, keycol, valuecol, gathercols)
```

Optional: Rename the factor levels of the variable column, and sort.

```{r}
# Rename factor names from "cond1" and "cond2" to "first" and "second"
levels(data_long$condition)[levels(data_long$condition)=="cond1"] <- "first"
levels(data_long$condition)[levels(data_long$condition)=="cond2"] <- "second"

# Sort by subject first, then by condition
data_long <- data_long[order(data_long$subject, data_long$condition), ]
data_long
```

### 4.2 From long to wide
Use ```spread```:  
```{r}
olddata_long
library(tidyr)

# The arguments to spread():
# - data: Data object
# - key: Name of column containing the new column names
# - value: Name of column containing values
data_wide <- spread(olddata_long, condition, measurement)
data_wide
```

Optional: A few things to make the data look nicer.  
```{r}
# Rename cond1 to first, and cond2 to second
names(data_wide)[names(data_wide)=="cond1"] <- "first"
names(data_wide)[names(data_wide)=="cond2"] <- "second"

# Reorder the columns
data_wide <- data_wide[, c(1,2,5,3,4)]
data_wide
```

The order of factor levels determines the order of the columns. The level order can be changed before reshaping, or the columns can be re-ordered afterward.

# 5.reshape2

### 5.1 From wide to long
Use ```melt```:  
```{r}
olddata_wide
library(reshape2)

# Specify id.vars: the variables to keep but not split apart on
melt(olddata_wide, id.vars=c("subject", "sex"))
```

There are options for melt that can make the output a little easier to work with:  

```{r}
data_long <- melt(olddata_wide,
        # ID variables - all the variables to keep but not split apart on
    id.vars=c("subject", "sex"),
        # The source columns
    measure.vars=c("control", "cond1", "cond2" ),
        # Name of the destination column that will identify the original
        # column that the measurement came from
    variable.name="condition",
    value.name="measurement"
)
data_long
```

If you leave out the measure.vars, melt will automatically use all the other variables as the id.vars. The reverse is true if you leave out id.vars.  

If you don’t specify variable.name, it will name that column "variable", and if you leave out value.name, it will name that column "measurement".  

Optional: Rename the factor levels of the variable column.   

```{r}
# Rename factor names from "cond1" and "cond2" to "first" and "second"
levels(data_long$condition)[levels(data_long$condition)=="cond1"] <- "first"
levels(data_long$condition)[levels(data_long$condition)=="cond2"] <- "second"

# Sort by subject first, then by condition
data_long <- data_long[ order(data_long$subject, data_long$condition), ]
data_long
```

### 5.2 From long to wide
***
The following code uses dcast to reshape the data. This function is meant for data frames; if you are working with arrays or matrices, use acast instead.  

```{r}
olddata_long
# From the source:
# "subject" and "sex" are columns we want to keep the same
# "condition" is the column that contains the names of the new column to put things in
# "measurement" holds the measurements
library(reshape2)

data_wide <- dcast(olddata_long, subject + sex ~ condition, value.var="measurement")
data_wide
```

Optional: A few things to make the data look nicer.

```{r}
# Rename cond1 to first, and cond2 to second
names(data_wide)[names(data_wide)=="cond1"] <- "first"
names(data_wide)[names(data_wide)=="cond2"] <- "second"

# Reorder the columns
data_wide <- data_wide[, c(1,2,5,3,4)]
data_wide

```

